---
title: "Vectors in R"
author: "Samanyu Goyal"
date: "3/14/2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Vectors in R  

### 1. Introduction 

In R, a **vector** is used to display a collection of values composing of the same data type.
For example, a vector can contain the following:

- **Numeric** (Decimals + Integers)
- **Characters** (Strings + Letters)
- **Logical** (True or False)

Every value inside a vector has a given length of 1. the disctintion between **integers** and **numeric** values is quite small; therefore, R converts between them automatically. To keep things simple, both are considered **numeric**

## 2. How to Create a Vector

Creating a vector is a simple yet powerful process. Follow the steps outlined below:

**- Decide the type of vector you want to create.**
    As we mentioned earlier, there are multiple types of vectors and you want to start thinking about which one you will create.
    
**- Use the correct syntax**
    There are two correct syntax when making a vector. 
        variable_vector = c(1,2,3,4,5)
        variable_vector <- c(1,2,3,4,5)
    The two syntax that can use are **=** or **<-**
      
**- Use the c() function to combine multiple values and create a vector**

#### Examples  

```{r}
# Numeric vector --> The below is a numeric vector as it consists of decimal points
numeric_vector = c(1.2, 3.4, 5.6)

# Integer vector --> The L specifies the number as an integer rather than a numeric value
integer_vector = c(1L, 2L, 3L)

# Character vector
char_vec = c("apple", "banana")

# Logical vector
logical_vec = c(TRUE, FALSE, TRUE)

# Scalar as a vector of length 1
scalar <- 33

```
You can use typeof( ) to check the type of vector

```{r}
typeof(integer_vector)
```

## 3. How to Access Vectors


**a. Indexing inside a vector starts from 1. This means will return the first term in the vector**
```{r}
numeric_vector[1]
```


**b. Imagine having a long vector. How do we find the total length of it? We can use:**
```{r}
length(numeric_vector)
```

## 4. Some Fun Vector Operations

**a. Minimum and Maximum**
```{r}
#The value below is of your entire vector
print((numeric_vector))
#The min() command has displayed the smallest value
print(paste('min = ', min(numeric_vector)))
#The max() command has displayed the largest value
print(paste('max = ', max(numeric_vector)))
```

**b. Sum**
```{r}
vector1 <- c(1,2,3,4,5)
vector2 <- c(1,2,3,4,5)
#The command below will allow us to sum the vectors. 
#The first index will add with each other, same with the second and so on.
vector1 + vector2
```
**c. Order**
```{r}
vector_order <- c(1, 2, 8, 4, 5)

# Get the order of indexes that will sort the vector
ordered_indices <- order(vector_order)

# Print the ordered indices
print(ordered_indices)

# Use the ordered indices to reorder the vector
sorted_vector <- vector_order[ordered_indices]

# Print the sorted vector
print(sorted_vector)

```
## 5. Real World Examples
If you guys have learned vectors in math class, you understand the pain it is to find the angle between the two. Imagine this problem

Vector A goes through (1,2) and Vector B goes through (3,4). Assume both start at (0,0). Find Î¸ between the vectors. 

**Below is a great example of how we can solve it**
```{r}
#Here are our two vectors that we are trying to solve
vectora <- c(1,2)
vectorb <- c(3,4)

# The formula is cos(Î¸) = (a*b)/(||a||)(||b||)

#We need to get the access to (a*b)
top_vector <- (vectora[1]*vectorb[1]) + (vectora[2]*vectorb[2])

#These two statements help us find the magnitude
a_magnitude <- sqrt((vectora[1]*vectora[1]) + (vectora[2]*vectora[2]))
b_magnitude <- sqrt((vectorb[1]*vectorb[1]) + (vectorb[2]*vectorb[2]))

#Formula to find cos(Î¸)
cos_theta <- (top_vector)/(a_magnitude*b_magnitude)
theta <- (acos(cos_theta))

#Just so we can have the degrees!!!ðŸ˜
theta_degrees <- theta * (180/pi)
print(theta_degrees)
```

